---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pointblankops

<!-- badges: start -->
<!-- badges: end -->

The goal of pointblankops is to provide specialized data validation operations using lightweight operatives for focused intelligence gathering. It offers memory-efficient alternatives to pointblank agents for failure detection and reporting.

## Installation

You can install the development version of pointblankops from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("petrbouchal/pointblankops")
```

## Example

This is a basic example which shows you how to use operatives for data validation:

```{r}
#| label: example
#| eval: false
library(pointblankops)
library(dplyr)

# Create some test data
test_data <- data.frame(
  batch = c("A", "A", "B", "B", "C"),
  id = c(1, 2, 3, 4, 5),
  value = c(10, NA, 15, 8, 12),
  category = c("X", "Y", "X", "Z", "Y")
)

# Create an operative and add validation steps
operative <- create_operative(test_data) %>%
  pointblank::col_vals_not_null(columns = dplyr::vars(value)) %>%
  pointblank::col_vals_between(columns = dplyr::vars(value), left = 5, right = 20)

# Debrief the operative to get only the failures
failures <- debrief(operative, row_id_col = c("batch", "id"))

print(failures)

# For database operations, install DBI package:
# install.packages("DBI")

# For parquet file operations, install arrow package:
# install.packages("arrow")
```

## Key Features

- **Lightweight operatives**: Streamlined alternatives to pointblank agents
- **Memory-efficient processing**: Chunked processing for large datasets  
- **Multiple output formats**: Return tibbles, save to parquet files, or write to databases
- **Database compatibility**: Works with local data frames and database tables (DuckDB, SQLite)
- **Flexible ID columns**: Support for multiple row identifier columns
- **Whimsical naming**: Follows pointblank's playful terminology (agents → operatives, interrogate → debrief)
